### 2011-05-05 paper/paper.tex
added code to implementation details

To test how these algorithms performed in practice, we implemented five algorithms for sampling: ball-walking, adaptive ball-walking, Gaussian-walking, adaptive Gaussian-walking, and hit-and-run.

@@

>>>>

To test how these algorithms performed in practice, we implemented five algorithms for sampling: ball-walking, adaptive ball-walking, Gaussian-walking, adaptive Gaussian-walking, and hit-and-run. The five methods were implemented in C++ and tested on several test regions of various dimensions: a cube, a sphere, and a regular simplex, all definied by separation oracles. Real numbers were represented by 64-bit double precision, although random numbers were generated with 31 bits.

We had to be able to sample from Gaussians (for Gaussian-walking) and we had to be able to uniformly sample directions in $\mathbb{R}^n$ (for ball-walking). To sample Gaussians, we use the following method from \cite{Marsaglia}: generate a random pair $(x,y)$ from the unit circle (sample from the square $[-1,1]^2$ until we get a point in the circle). Conditioned on $s = x^2 + y^2 < 1$, we get that
\begin{center}
$\displaystyle x_1 = x\sqrt{\frac{-\ln(s)}{s}}$ and $\displaystyle y_1 = y\sqrt{\frac{-\ln(s)}{s}}$
\end{center}
are independent Gaussian variables with standard deviation $\sigma = 1/\sqrt{2}$ centered at $0$. Once we can sample from Gaussians, it is easy to sample directions uniformly: we just choose every coordinate independently from a Gaussian of center $0$ with some standard deviation $\sigma$. It is easy to see that this is uniform over all directions. Let $P$ be the probaility distribution for this. For any $x\in\mathbb{R}^n$, we get that $P(x) = e^{-||x||_2^2 / 2\sigma^2}$, which depends only on the norm $||x||_2$, not on the direction.

Here is the code for the separation oracles which define the cube, the sphere, and the simplex:

\begin{quote}
\begin{verbatim}
bool cube_separation_oracle(point p) {
    for(int i = 0; i<p.n; i++) {
        if(p.x[i] <= -1 || p.x[i] >= 1) return false;
    }
    return true;
}

bool sphere_separation_oracle(point p) {
    real square_norm = 0.0;
    for(int i = 0; i<p.n; i++) {
        square_norm += p.x[i] * p.x[i];
    }
    return square_norm < 1.0;
}

bool simplex_separation_oracle(point p) {
    real c = (1.0 - sqrt(1.0 + (real)p.n)) / (real)p.n;
    real coordsum = 0.0;
    for(int i = 0; i < p.n; i++) {
         coordsum += p.x[i];
    }
    real lambda = (1.0 - (real)p.n * c) / (coordsum - (real)p.n * c);
    if(lambda < 1.0)
        return false;
    for(int i = 0; i < p.n; i++) {
        if(c + lambda * (p.x[i] - c) < 0.0f)
            return false;
    }
    return true;
}
\end{verbatim}
\end{quote}

@@

\bibitem{Marsaglia} Marsaglia, G. and Bray, T.A. ``A Convenient Method for Generating Normal Variables." \emph{SIAM Review} Vol. 6, No. 3 (Jul., 1964), pp. 260-264.
